---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ravetools

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/ravetools)](https://CRAN.R-project.org/package=ravetools)
[![R-CMD-check](https://github.com/dipterix/ravetools/workflows/R-CMD-check/badge.svg)](https://github.com/dipterix/ravetools/actions)
<!-- badges: end -->

The goal of `ravetools` is to provide memory-efficient signal processing toolbox for `intracranial-EEG` analyses. Highlighted features include:

* [`Notch filter`](#) (remove electric line noise)
* [`Welch Periodogram`](#) (averaged power over frequencies)
* [`Wavelet`](#) (frequency-time decomposition)

## Installation

The package is available on `CRAN`. To install the compiled version, simply run:

```r
install.packages("ravetools")
```

If you want to participate in the development of `ravetools`. Please make sure you have [`FFTW3`](https://www.fftw.org/) installed. On Windows, install [`Rtools`](https://cran.r-project.org/bin/windows/Rtools/) and add it to system path; on MacOS, install [`HomeBrew`](https://docs.brew.sh/Installation) in your shell terminals, and run `brew install pkg-config fftw`; on Ubuntu-Linux, run bash command `sudo apt-get install libfftw3-dev`. Please search online about how to install `FFTW3` on other operating systems. Once `FFTW3` is installed, execute the following R command.

``` r
#!/usr/bin/env R
# install.packages("remotes")
remotes::install_github("dipterix/ravetools")
```

## `iEEG` preprocess pipeline

This is a basic example which shows you how to preprocess an `iEEG` signal. The goal here is to:

* Plot diagnostic graphs to inspect channels
* Apply Notch filters to remove electric line noise
* Frequency-time decomposition and show the power densities

<small>* Channel referencing is not included</small>

### 1. Generate toy examples:

```{r}
library(ravetools)

# Generate 20 second data at 2000 Hz
time <- seq(0, 20, by = 1 / 2000)
signal <- sin( 120 * pi * time) + 
  sin(time * 20*pi) + 
  exp(-time^2) * 
  cos(time * 10*pi) +
  rnorm(length(time))

diagnose_channel(signal, srate = 2000)
```


### 2. Apply `Notch` filters and inspect `Periodograms`

```{r}
## ------- Notch filter --------
signal2 <- notch_filter(signal, sample_rate = 2000)

diagnose_channel(signal, signal2, srate = 2000,
                 name = c("Raw", "Filtered"))

```

### 3. Frequency-time decomposition 

Current version of `ravetools` provides two approaches: `Wavelet` and `Multi-taper`. `Wavelet` uses the [`Morlet wavelet`](https://en.wikipedia.org/wiki/Morlet_wavelet) and obtains both amplitude and phase data, while `Multi-taper` does not generate phase data. However, the amplitude obtained from `Multi-taper` is smoother than `Wavelet`. 

#### Using `Wavelet`:

```{r}
## ---------- Wavelet -----------
coef <- morlet_wavelet(
  signal2, freqs = seq(1, 100), 
  srate = 2000, wave_num = c(2, 15))
amplitude <- 10 * log10(Mod(coef[]))

# For each frequency, decimate to 100 Hz
downsample_amp <- apply(amplitude, 2, decimate, q = 20)
downsample_time <- decimate(time, q = 20)

par(mfrow = c(1,1))
image(
  z = downsample_amp,
  x = downsample_time,
  y = 1:100,
  xlab = "Time (s)",
  ylab = "Frequency (Hz)",
  main = "Amplitude (dB)",
  sub = "Wavelet at 2000 Hz, then down-sampled to 100 Hz", 
  col = matlab_palette()
)
```


#### `Multi-taper`

Alternatively you can use `Multi-tapers` to obtain amplitude data. The algorithm is modified from source code [here](https://github.com/preraulab/multitaper_toolbox). Please credit them as well if you adopt this approach.

```{r}
## ---------- Multitaper -----------
res <- multitaper(
  data = signal2,
  fs = 2000,
  frequency_range = c(1, 100),
  time_bandwidth = 1.5,
  window_params = c(2, 0.01)
)

par(mfrow = c(1,1))
image(
  x = res$time,
  y = res$frequency,
  z = 10 * log10(res$spec),
  xlab = "Time (s)",
  ylab = 'Frequency (Hz)',
  col = matlab_palette(),
  main = "Amplitude (dB)"
)
```

